---
title: "DMtest"
author: "James Dai and Xiaoyu Wang"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DMtest}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Introduction
Cancer development is ubiquitously associated with aberrant DNA methylation.
Noninvasive biomarkers incorporating DNA methylation signatures are of rising
interest for cancer early detection. Statistical tests for discovering dierential
methylation markers have largely focused on assessing dierences of mean
methylation levels, e.g., between cancer and normal samples. Previously two standard tests 
were applied for the task: differential methylated CpG (DMC) and
differential variability CpGs (DVC). In the package we propose tow joint tests (DMVC) 
combing DMC and DVC. The first DMVC is proposed to test hypermethylation and 
increased variance in cancer, and the second DMVC is aimed to test for differential methylation
in both direction and increased variance in cancer.

The proposed joint tests substantially improved detection power in 
simulation studies, and yields more cancer CpG markers than the standard tests.


# Example
The following example takes DNA methylation data from 334 samples of TCGA-COAD;
In the illustration we use top 1000 CpG probes to save time. For genome-wide data,
users can set appropriate corenumber and use the parallel version for the joint tests.

```{r setup}
library(DMtest)
#load example data
data(beta)
dim(beta)
data("covariate")
dim(covariate)
#compute p-values 
out=dmvc(beta=beta,covariate=covariate)
head(out)
```

# Reference
James Y. Dai et al. Incorporating increased variability in differential 
methylation and cancer biomarker discovery, 
Genome Biology 2021, submitted.